---
title: "Main results analysis"
output:
  html_document:
    css: custom_notebook_formatting.css
    toc: true
    toc_depth: 3
    df_print: paged
    code_folding: hide
---

## Preparations

```{r PREPS, message=FALSE, warning=FALSE}
library(ggplot2)
library(cowplot)
library(readr)
library(stringr)
library(dplyr)
library(kableExtra)
library(afex)

options(scipen = 100)
knitr::opts_chunk$set(results = "hold", fig.align = 'center')
source('src/analyses/plotting_functions.R')
#source('src/ml_pipeline/preprocessing.R')
```

## Load results

Load data from previously organized rdata file.

```{r LOAD_REUSLTS, results='hold'}
all_rdata_path <- 'data/ml_output/results_tables/results_for_analysis.RData'
load(all_rdata_path)
rm(all_rdata_path)
```

Change to "pretty" dataset names:

```{r}
new_dataset_names <-c(
  "cd_franzosa_2019"="CD (Franzosa, 2019)",
  "cirrhosis_qin_2014"="Cirrhosis (Qin, 2014)",
  "crc_feng_2015"="CRC (Feng, 2015)",
  "crc_s3_s4_yachida_2019"="CRC (Yachida, 2019)",
  "crc_yu_2015"="CRC (Yu, 2015)",
  "esrd_wang_2020"="ESRD (Wang, 2020)",
  "metacardis_1_8"="MS (MetaCardis)",
  "metacardis_3_8"="T2D (MetaCardis)",
  "sth_rubel_2020"="STH (Rubel, 2020)",
  "uc_franzosa_2019"="UC (Franzosa, 2019)",
  "uc_spain_nielsen_2014"="UC (Nielsen, 2014)"
  )

cv_datasets_summary$dataset <- new_dataset_names[cv_datasets_summary$dataset]
cv_results$dataset <- new_dataset_names[cv_results$dataset]
feat_imp$dataset <- new_dataset_names[feat_imp$dataset]
feat_imp_sum$dataset <- new_dataset_names[feat_imp_sum$dataset]
```


### Descriptive statistics

```{r FIRST_CV_STATS, results='hold'}
cv_datasets_summary %>% 
  filter(!shuffled) %>%
  select(-shuffled, -tuned, -feature_set_type, -fs_type) %>%
  kbl(digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 
```

### More settings

```{r}
# datasets_to_focus_on <- c(
#   'cd_franzosa_2019',
#   'uc_franzosa_2019',
#   'crc_s3_s4_yachida_2019',
#   'cirrhosis_qin_2014'
# )
```

```{r}
feature_type_color_map = 
  c("T" = "firebrick4", 
    "S" = "chartreuse4", 
    "P" = "darkorange3", 
    "M" = "midnightblue", 
    "All" = "grey30")

# Save number of folds/repeats
n_folds <- n_distinct(cv_results$fold_id)
```

## Supplementary Table S2: Random forest model performance results using early integration

```{r}
cv_datasets_summary %>%
  filter(!shuffled) %>%
  select(dataset, feature_set_type, fs_type, 
         n_features_origin_T, n_features_origin_P, n_features_origin_M, n_features_origin_S,
         mean_out_of_fold_test_auc) %>%
  mutate(mean_out_of_fold_test_auc = round(mean_out_of_fold_test_auc, 3)) %>%
  rename(Dataset = dataset) %>%
  rename(`Features used` = feature_set_type) %>%
  rename(`Feature selection method` = fs_type) %>%
  rename(`Number of original features - T (Taxa)` = n_features_origin_T) %>%
  rename(`Number of original features - S (Serum metabolites)` = n_features_origin_S) %>%
  rename(`Number of original features - P (Pathways)` = n_features_origin_P) %>%
  rename(`Number of original features - M (Fecal metabolites)` = n_features_origin_M) %>%
  rename(`Cross-validated average AUC` = mean_out_of_fold_test_auc) %>%
  write_csv("docs/Table_S2.csv")
```

## Plot: Number of samples and number of features

A. Number of samples per study. C, controls (dark grey); D, disease (light grey). 
B. Average number of original features from each type. 
(Or: number of features selected by the feature-selection process and used for final training. The average is over the cross-validation process).

```{r, fig.height=5.1, fig.width=6, warning = FALSE}
plot_basic_stats(
  cv_results = cv_results, 
  feature_type_color_map = feature_type_color_map
  )
```

## Feature importance analysis

### Supplementary Table S3

Only provided for models with an AUC > 0.7, and contributing features.

```{r}
feat_imp_sum %>%
  filter(mean_out_of_fold_test_auc > 0.7) %>%
  select(dataset, feature_set_type, feature, 
         n_times_selected, mean_importance, 
         combined_fdr, contributor) %>%
  mutate(mean_importance = round(mean_importance, 5)) %>%
  mutate(combined_fdr = ifelse(
    combined_fdr < 0.00001, '< 0.00001', 
    round(combined_fdr, 5))) %>%
  rename(Dataset = dataset) %>%
  rename(`Feature set` = feature_set_type) %>%
  rename(`Feature name` = feature) %>%
  rename(`RF permutation importance, averaged over folds` = mean_importance) %>%
  rename(`Altmann FDR-corrected p value, aggregated over folds using Fisher's method` = combined_fdr) %>%
  rename(`Number of times selected by feature selection (out of 100 repeats/folds)` = n_times_selected) %>%
  filter(contributor) %>%
  rename(`Overall contributor` = contributor) %>%
  mutate(`Feature description` = get_feature_descriptions(`Feature name`)) %>%
  write_csv("docs/Table_S3.csv")
```

## Session info

```{r}
sessionInfo()
```